Symbolic linearization of equations of motion of constrained multibody systems

### 2013-08-01 PetersonGedeHubbard2012.tex
Rewrote example for new speeds.
Now uses the lean-frame fixed speeds for the rolling disc example. Might need a
little cleanup, and probably the figure.

\begin{equation}
\label{eq:definition_Fstar}
\mathbf{F} =
\begin{bmatrix}
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 \\
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 \\
\displaystyle \vdots \\
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o
\end{bmatrix}
\end{equation}

@@

\section{Dynamics Example}
To illustrate Kane's method in the presence of configuration and velocity
constraints, we consider a thin disc $C$ rolling without slip on a horizontal
plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
with $\hat{c}_y$ perpendicular to the disc plane. We take $r$ and $m$ to be the
disc radius and mass, respectively, and assume the mass distribution of a thin
disc so that $I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. We take the inertial
(Newtonian) frame $N$ to be similarly equipped with dextral unit vectors
$\hat{n}_x, \hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground
plane and aligned with local gravitational field ($\hat{n}_z$ points downwards).
To orient the disc, we first align the $C$ with $N$, then apply successive body
fixed ZYX rotations: yaw $q_1$, roll $q_2$, and spin $q_3$.  We denote by $A$
and $B$ the two frames associated with the first two intermediate rotations in
the sequence; we refer to these as the instantaneous yaw and lean frames,
respectively. The unit vector $\hat{b}_z$ fixed in the $B$ frame is parallel to
the disc plane and aligned with the position vector from the disc center $C^*$
to the lowest point of the disc $P^*$. Traditionally, to locate the center of
the disc $C^*$ relative to the inertial origin $N^*$, two coordinates are used
to locate the contact point in the ground plane; those two coordinates plus the
lean angle $q_2$ determine the location of the center of the disc. The
advantage of this approach is that the choice of coordinates guarantees the
disc remain in contact with the ground plane. To illustrate how dependent
coordinates must be accounted for when linearizing the equations of motion, we
purposefully locate the center of the disc $C^*$ relative to the inertial
origin $N^*$ with a non-minimal choice of coordinates
\begin{equation*}
  \bar{r}^{C^*} = q_4 \hat{n}_x + q_5 \hat{n}_y + q_6 \hat{n}_z
\end{equation*}
which results in the configuration constraint
\begin{equation}
  r\cos{q_2} + q_6 = 0
\end{equation}
which must be satisfied for the disc to remain in contact with the ground.
While this constraint is easily avoided by appropriate choice of coordinates,
it serves to illustrate considerations which must be made when analyzing
systems where there is no obvious way to eliminate configuration constraints.

Again for the purpose of illustrating how velocity constraints must be
accounted for when obtaining linearized equations of motion, we define
the angular velocity of $C$ and the velocity of $C^*$ relative to $N$ using a
non-minimal choice of generalized speeds
\begin{align}
  \label{w_u}
  {^N}\bar{\omega}^C &= u_1 \hat{c}_x + u_2 \hat{c}_y + u_3 \hat{c}_z \\
  \label{v_u}
  {^N}\bar{v}^{C^*} &= u_4 \hat{c}_x + u_5 \hat{c}_y + u_6 \hat{c}_z
\end{align}
It should be immediately obvious from these defintions that the partial angular
velocity and partial velocity of $C^*$, respectively, are
\begin{align}
  {^N}\bar{\omega}^C_u &= [\hat{c}_x, \hat{c}_y, \hat{c}_z, 0, 0, 0] \\
  {^N}\bar{v}^{C^*}_u &= [0, 0, 0, \hat{c}_x, \hat{c}_y, \hat{c}_z]
\end{align}
where we present the partial velocities as a $1\times6$ matrix whose entries
are populated with standard 3-vectors. Given these definitions, the velocity of
the lowest point of the disc $P^*$ is
\begin{align*}
  \bm{v}^{P^*} &= \bm{v}^{C^*} + \bm{\omega}^{C} \times r^{P^*/C^*} \\
  &= (r u_2 \cos q_3  + u_4) \bm{c}_x
   + (-r (u_1 \cos q_3  + u_3 \sin q_3) + u_5) \bm{c}_y \\
   &+ (r u_2 \sin q_3  + u_6) \bm{c}_z
\end{align*}
where we have made use of the fact that $\bm{r}^{P^*/C^*} = r\bm{b}_z$.
Under the assumption of pure rolling this immediately yields three velocity
constraints
\begin{align}
  r u_2 \cos q_3  + u_4 &= 0\\
  -r (u_1 \cos q_3  + u_3 \sin q_3) + u_5 &= 0\\
  r u_2 \sin q_3  + u_6 &= 0
\end{align}
These equations are linear in the $u_i$'s, nonlinear in the $q_i$'s, involve
geometric system parameters (but not mass or inertial parameters), and do not
explicitly involve $\dot{q}_i$'s. This structure will be taken advantage of
subsequently.

At this point, the reader familiar with the rolling disc problem is certainly
wondering why we have chosen such a diabolical set of coordinates and speeds.
In this simple example, there is an \textit{obvious and minimal} choice of
coordinates and speeds ($q_6, u_4, u_5, u_6$, should be dependent; this choice
will be valid for all parameters and configurations; further, the derivation of
the equations of motion can be done without ever introducing these quantities),
for some systems this is not the case (e.g., Stewart's platform, bicycle). The
numeric values of system parameters and the configurations in which the system
will operate will determine which coordinates and speeds can be taken to be
dependent (and hence potentially eliminated by algebraic manipulations).
However, it may not be possible \textit{apriori} to easily determine which
coordinates and speeds should be taken as dependent; further, some systems may
operate in distinct enough regions of the configuration space that one
\textit{must} use different choices of dependent coordinates and/or speeds as
the system moves from one region to another. The purpose of our diabolical
choice of coordinates and speeds is to illustrate, in the context of a familiar
example, how constraints must be taken into account when linearizing equations
of motion. It is our hope that by doing this for a well known and relatively
simple example, readers can apply the same techniques to more complicated
systems where its use is more appropriate.

Time differentiating the velocity constraints yields
\begin{align}
-r u_2 \dot{q}_3 \sin q_3 + r \dot{u}_2 \cos q_3 + \dot{u}_4 &= 0 \\
r(u_1 \dot{q}_3 \sin q_3 - u_3 \dot{q}_3 \cos q_3) - r \dot{u}_1 \cos q_3 - r \dot{u}_3 \sin q_3 + \dot{u}_5 &= 0 \\
r u_2 \dot{q}_3 \cos q_3 + r \dot{u}_2 \sin q_3 + \dot{u}_6 &= 0
\end{align}
which must also be satisfied during general motions. There are two types of
terms in these equations; 1) terms involving $\dot{u}_i$'s and $q_i$'s, and 2)
terms involving $\dot{q}_i$'s, $u_i$'s, and $q_i$'s. This structure will be
taken advantage of subsequently.

The kinematic differential equations relate the time derivatives of the
coordinates and the generalized speeds. From the addition theorem for angular
velocity, we may write the angular velocity of $C$ as
\begin{equation}
  \label{w_qdot}
  {^N}\bar{\omega}^C = \dot{q}_1 \hat{n}_z + \dot{q}_2 \hat{a}_x + \dot{q}_3 \hat{b}_y
\end{equation}
while time differentiating $r^{C^*}$ in $N$ yields
\begin{equation}
  \label{v_qdot}
  \bar{v}^{C^*} = \dot{q}_4 \hat{n}_x + \dot{q}_5 \hat{n}_y + \dot{q}_6 \hat{n}_z
\end{equation}
Equating (\ref{w_u}) with (\ref{w_qdot}) and (\ref{v_u}) with (\ref{v_qdot}),
and resolving these vector equations into the $B$ frame, and rearranging so
that all terms appear to the left of the equality sign, we obtain
\begin{align}
  \dot{q}_2 - u_1 \cos q_3  - u_3 \sin q_3 &= 0 \\
  \dot{q}_1 \sin q_2 + \dot{q}_3 - u_2 &= 0 \\
  \dot{q}_1 \cos q_2 + u_1 \sin q_3  - u_3 \cos q_3  &= 0 \\
  \dot{q}_4 \cos q_1 + \dot{q}_5 \sin q_1 -u_4 \cos q_3 - u_6 \sin q_3 &= 0 \\
 -\dot{q}_4 \sin q_1 \cos q_2 + \dot{q}_5 \cos q_1 \cos q_2 + \dot{q}_6 \sin q_2 - u_5 &= 0 \\
  \dot{q}_4 \sin q_1 \sin q_2 - \dot{q}_5 \sin q_2 \cos q_1 + \dot{q}_6 \cos q_2 + u_4 \sin q_3 - u_6 \cos q_3 &= 0
\end{align}
The angular acceleration of $C$ and the translation acceleration of $C^*$,
relative to $N$, is
\begin{align}
  {^N}\bar{\alpha}^{C} &= \dot{u}_1 \hat{c}_x + \dot{u}_2 \hat{c}_y + \dot{u}_3 \hat{c}_z  \\
  {^N}\bar{a}^{C^*} &= (u_2 u_6 - u_3 u_5 + \dot{u}_4) \hat{c}_x + (u_3 u_4 - u_1 u_6 +
+ \dot{u}_5) \hat{c}_y \notag\\&+ (u_1 u_5 - u_2 u_4 + \dot{u}_6) \hat{c}_z
\end{align}

>>>>

\begin{equation}
\label{eq:definition_Fstar}
\mathbf{F}^* =
\begin{bmatrix}
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 \\
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 \\
\displaystyle \vdots \\
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o
\end{bmatrix}
\end{equation}

@@

\section{Dynamics Example}
To illustrate Kane's method in the presence of configuration and velocity
constraints, we consider a thin disc $C$ rolling without slip on a horizontal
plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
with $\hat{c}_y$ perpendicular to the disc plane. We take $r$ and $m$ as the
disc radius and mass, and assume a thin disc mass distribution so that
$I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. We take the inertial (Newtonian)
frame $N$ to be similarly equipped with dextral unit vectors $\hat{n}_x,
\hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground plane
(downwards) and aligned with local gravitational field. To orient the disc, we
first align the $C$ with $N$, then apply successive body fixed ZXY rotations:
yaw $q_1$, roll $q_2$, and spin $q_3$.  We denote by $A$ and $B$ the two frames
associated with the first two intermediate rotations in the sequence; we refer
to these as the instantaneous yaw and lean frames, respectively. The unit
vector $\hat{b}_z$ fixed in the $B$ frame is parallel to the disc plane and
aligned with the position vector from the disc center $C^*$ to the lowest point
of the disc $P^*$. Traditionally, to locate the center of the disc $C^*$
relative to the inertial origin $N^*$, two coordinates are used to locate the
contact point in the ground plane; those two coordinates plus the lean angle
$q_2$ determine the location of the center of the disc. The advantage of this
approach is that the choice of coordinates guarantees the disc remain in
contact with the ground plane. To illustrate how dependent coordinates must be
accounted for when linearizing the equations of motion, we purposefully locate
the center of the disc $C^*$ relative to the inertial origin $N^*$ with a
non-minimal choice of coordinates
\begin{equation*}
  \bar{r}^{N^*C^*} = q_4 \hat{n}_x + q_5 \hat{n}_y + q_6 \hat{n}_z
\end{equation*}
which results in the configuration constraint $\mathbf{f}_c$
\begin{equation}
  \label{rd:f_c}
  r\cos{q_2} + q_6 = 0
\end{equation}
which must be satisfied for the disc to remain in contact with the ground.
While this constraint is easily avoided by appropriate choice of coordinates,
it serves to illustrate considerations which must be made when analyzing
systems where there is no obvious way to eliminate configuration constraints.

Again to demonstrate how velocity constraints must be accounted for when
obtaining linearized equations of motion, we define the velocity of $C^*$ in
$N$ with a non-minimal choice of speeds, defined in the body-fixed ($C$) frame.
The angular velocity of $C$ in $N$, however, is defined in the intermediate
frame $B$, which is the instantaneous lean frame.
\begin{align}
  \label{v_u}
  {^N}\bar{v}^{C^*} &= u_4 \hat{c}_x + u_5 \hat{c}_y + u_6 \hat{c}_z \\
  \label{w_u}
  {^N}\bar{\omega}^C &= u_1 \hat{b}_x + u_2 \hat{b}_y + u_3 \hat{b}_z
\end{align}
It should be immediately obvious from these defintions that the partial angular
velocity and partial velocity of $C^*$, respectively, are
\begin{align}
  {^N}\bar{v}^{C^*}_u &= [0, 0, 0, \hat{c}_x, \hat{c}_y, \hat{c}_z] \\
  {^N}\bar{\omega}^C_u &= [\hat{b}_x, \hat{b}_y, \hat{b}_z, 0, 0, 0] 
\end{align}
where we present the partial velocities as a $1\times6$ matrix whose entries
are populated with standard 3-vectors. Given these definitions, the velocity of
the lowest point of the disc $P^*$ is
\begin{align*}
    {^N}\bm{v}^{P^*} &= \bm{v}^{C^*} + \bm{\omega}^{C} \times r^{P^*/C^*} \\
                     &= u_4\hat{c}_x + u_5\hat{c}_y + u_6\hat{c}_z + r u_2
                        \hat{b}_x - r u_1 \hat{b}_y \\
                     &= (r u_2 \cos(q_3) + u_4) \hat{c}_x + (-r
                        u_1 + u_5)\hat{c}_y + r u_2 \sin(q_3) +
                        u6) \hat{c}_z
\end{align*}
where we have made use of the fact that $\bm{r}^{P^*/C^*} = r\bm{b}_z$.
Under the assumption of pure rolling this immediately yields three velocity
constraint equations which make up $\mathbf{f}_v = \mathbf{0}$
\begin{subequations}
\label{rd:f_v}
\begin{align}
    r u_2 \cos(q_3) + u_4 &= 0\\
            -r u_1 + u_5 &= 0\\
    r u_2 \sin(q_3) + u_6 &= 0
\end{align}
\end{subequations}
These equations are linear in the $u_i$ terms, nonlinear in the $q_i$ terms,
involve geometric system parameters (but not mass or inertial parameters), and
do not explicitly involve $\dot{q}_i$ terms. This structure will be taken
advantage of subsequently.

At this point, the reader familiar with the rolling disc problem is certainly
wondering why we have chosen such a diabolical set of coordinates and speeds.
In this simple example, there is an {\bf \textit{obvious and minimal}} choice of
coordinates and speeds ($q_6, u_4, u_5, u_6$, should be dependent; this choice
will be valid for all parameters and configurations; further, the derivation of
the equations of motion can be done without ever introducing these quantities),
for some systems this is not the case (e.g., Stewart's platform, bicycle). The
numeric values of system parameters and the configurations in which the system
will operate will determine which coordinates and speeds can be taken to be
dependent (and hence potentially eliminated by algebraic manipulations).
However, it may not be possible \textit{apriori} to easily determine which
coordinates and speeds should be taken as dependent; further, some systems may
operate in distinct enough regions of the configuration space that one
\textit{must} use different choices of dependent coordinates and/or speeds as
the system moves from one region to another. The purpose of our diabolical
choice of coordinates and speeds is to illustrate, in the context of a familiar
example, how constraints must be taken into account when linearizing equations
of motion. It is our hope that by doing this for a well known and relatively
simple example, readers can apply the same techniques to more complicated
systems where its use is more appropriate or necessary.

Time differentiating the velocity constraint equations yields the acceleration
constraint equations, $\mathbf{f}_a = \mathbf{0}$, written out as:
\begin{subequations}
\label{rd:f_a}
\begin{align}
    -r u_{2} \sin(q_{3}) \dot{q}_{3} + r \cos(q_{3}) \dot{u}_{2} + \dot{u}_{4} &=
    0\\
    - r \dot{u}_{1} + \dot{u}_{5} &= 0\\
    r u_{2} \cos(q_{3}) \dot{q}_{3} + r \sin(q_{3}) \dot{u}_{2} + \dot{u}_{6} &=
    0
\end{align}
\end{subequations}
which must also be satisfied during general motions. There are two types of
terms in these equations; 1) terms involving $\dot{u}_i$'s and $q_i$'s, and 2)
terms involving $\dot{q}_i$'s, $u_i$'s, and $q_i$'s. This structure will be
taken advantage of subsequently.

The kinematic differential equations relate the time derivatives of the
coordinates and the generalized speeds. Here they are solved by equating the
velocities written with time-differentiated generalized coordinates and the
velocities written with generalized speeds.  From the addition theorem for
angular velocity, we may write the angular velocity of $C$ as
\begin{equation}
  \label{w_qdot}
  {^N}\bar{\omega}^C = \dot{q}_1 \hat{n}_z + \dot{q}_2 \hat{a}_x + \dot{q}_3 \hat{b}_y
\end{equation}
while time differentiating $r^{C^*}$ in $N$ yields
\begin{equation}
  \label{v_qdot}
  \bar{v}^{C^*} = \dot{q}_4 \hat{n}_x + \dot{q}_5 \hat{n}_y + \dot{q}_6 \hat{n}_z
\end{equation}
Equating (\ref{w_u}) with (\ref{w_qdot}) and (\ref{v_u}) with (\ref{v_qdot}),
and resolving these vector equations into the $B$ frame, and rearranging so
that all terms appear to the left of the equality sign, we obtain
\begin{align}
    \label{rd:f_0_f_1}
\underbrace{\left[\begin{matrix}\dot{q}_{2}\\\sin\left(q_{2}\right)
    \dot{q}_{1} + \dot{q}_{3}\\\cos\left(q_{2}\right)
    \dot{q}_{1}\\\dot{q}_{4}\\\dot{q}_{5}\\\dot{q}_{6}\end{matrix}\right]}_{\mathbf{f}_0}
    + 
\underbrace{\left[\begin{matrix}- u_{1}\\- u_{2}\\- u_{3}\\r u_{1}
    \sin\left(q_{1}\right) \cos\left(q_{2}\right) +
    r u_{2} \cos\left(q_{1}\right)\\- r u_{1}
    \cos\left(q_{1}\right) \cos\left(q_{2}\right) +
    r u_{2} \sin\left(q_{1}\right)\\- r u_{1}
    \sin\left(q_{2}\right)\end{matrix}\right]}_{\mathbf{f}_1}
    = \left[\begin{matrix} 0\\ 0\\ 0\\ 0\\ 0\\ 0\end{matrix}\right]
\end{align}
The translation acceleration of $C^*$ and the angular acceleration of $C$,
relative to $N$, are
\begin{align}
    {^N}\bar{a}^{C^*} =& (-(u_1 \sin(q_3) + u_3 \cos(q_3)) u_5 + u_2 u_6 +
                         \dot{u}_4) \hat{c}_x \notag \\
                      & + ((u_1 \sin(q_3) + u_3 \cos(q_3)) u_4 - (u_1
                         \cos(q_3) - u_3 \sin(q_3)) u_6 + \dot{u}_5) \hat{c}_y
                         \notag\\
                      & + ((u_1 \cos(q_3) - u_3 \sin(q_3)) u_5 - u_2 u_4 +
                         \dot{u}_6) \hat{c}_z \\
    {^N}\bar{\alpha}^{C} =& (-u_2 u_3 + u_3^2 \tan(q_2) + \dot{u}_1) \hat{b}_x
                         + \dot{u}_2 \hat{b}_y + (u_1 u_2 - u_1 u_3 \tan(q2) +
                         \dot{u}_3) \hat{b}_z
\end{align}

As stated, the mass of the disc is $m$, and the inertia dyadic of the disc can
be written as
\begin{align}
    \bar{\bar{\mathbf{I}}}^{C/C^*} = \frac{m r^2}{4} \hat{c}_x\hat{c}_x +
    \frac{m r^2}{2} \hat{c}_y\hat{c}_y + \frac{m r^2}{4} \hat{c}_z\hat{c}_z
\end{align}
Using equations (\ref{eq:rb_translational_gen_inertia}) and
(\ref{eq:rb_rotational_gen_inertia}), $\bar{R}^*_{C^*}$ and $\bar{T}^*_C$ can be
written.
\begin{align}
    \bar{R}^*_{C^*} =& - m (- (u_{1} \sin(q_{3}) + u_{3}
                       \cos(q_{3})) u_{5} + u_{2} u_{6} +
                       \dot{u}_{4}) \hat{c}_x \notag \\
                     & - m ((u_{1} \sin(q_{3}) + u_{3}
                       \cos(q_{3})) u_{4} - (u_{1}
                       \cos(q_{3}) - u_{3}
                       \sin(q_{3})) u_{6} + \dot{u}_{5})
                       \hat{c}_y \notag \\
                     & - m ((u_{1} \cos(q_{3}) - u_{3}
                       \sin(q_{3})) u_{5} - u_{2} u_{4} +
                   \dot{u}_{6}) \hat{c}_z \\
    \bar{T}^*_C =& \frac{m r^{2}}{4} (2 u_{1} u_{2} \sin(q_{3})
                   - u_{1} u_{3} \sin(q_{3}) \tan(q_{2})
                   + 2 u_{2} u_{3} \cos(q_{3}) \notag\\&\quad\quad- u_{3}^2
                   \cos(q_{3}) \tan(q_{2}) +
                   \sin(q_{3}) \dot{u}_{3} - \cos(q_{3})
                   \dot{u}_{1}) \hat{c}_x \notag \\
                 & - \frac{m r^{2}}{2} \dot{u}_{2} \hat{c}_y \notag \\
                 & + \frac{m r^{2}}{4} (- 2 u_{1} u_{2}
                   \cos(q_{3}) + u_{1} u_{3} \cos(q_{3})
                   \tan(q_{2}) + 2 u_{2} u_{3}
                   \sin(q_{3}) \notag\\&\quad\quad - u_{3}^2 \sin(q_{3})
                   \tan(q_{2}) - \sin(q_{3}) \dot{u}_{1}
                   - \cos(q_{3}) \dot{u}_{3}) \hat{c}_z
\end{align}

The generalized active forces for this example are
\begin{align}
    \bar{R}_{C^*} &= m g \hat{n}_z \\
    \bar{T}_C &= 0
\end{align}

Equations (\ref{eq:definition_F}), (\ref{eq:definition_Fstar}), and
(\ref{eq:kanes_eq}) can be used to construct the equations of motion.
For brevity, we omit the results of each step involved, and only write the
results, after they have been partitioned into the $\mathbf{f}_2$ and
$\mathbf{f}_3$ vectors.
\begin{align}
    \label{rd:f_2_f_3}
    & \left[\begin{matrix} 0\\ 0\\ 0\end{matrix}\right] =
    \underbrace{\left[\begin{matrix}- \frac{m r}{4} \left(r \dot{u}_{1} + 4
                \dot{u}_{5}\right)\\\frac{m r}{2} \left(- r \dot{u}_{2} + 2
                \sin\left(q_{3}\right) \dot{u}_{6} + 2 \cos\left(q_{3}\right)
                \dot{u}_{4}\right)\\- \frac{m r^{2}}{4}
                \dot{u}_{3}\end{matrix}\right]}_{\mathbf{f}_2} + \notag\\
    &\underbrace{\left[\begin{matrix}
                m r \left( g \sin(q_2) - u_1 u_4 \sin(q_3) + u_1 u_6 \cos(q_3)
                    + (\frac{r u_2}{2} - \frac{r u_3}{4} \tan(q_2) - u_4
                \cos(q_3) - u_6 \sin(q_3)) u_3 \right)
                \\m r \left(- u_{2} u_{4}
                \sin\left(q_{3}\right) + u_{2} u_{6}
                \cos\left(q_{3}\right) - u_{3} u_{5}\right)\\\frac{m r^{2}}{4}
                \left(- 2 u_{2} + u_{3} \tan\left(q_{2}\right)\right)
                u_{1}\end{matrix}\right]}_{\mathbf{f}_3} 
\end{align}

As there are 12 unknown quantities, $\dot{q}_{1-6}$ and $\dot{u}_{1-6}$, we
need 3 additional equations so solve for the time-derivative terms.
The acceleration constraints $\mathbf{f}_a$ provide the final 3 equations.
Combining (\ref{rd:f_0_f_1}), (\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the
complete second order differential equations can be written.
\begin{align}
\label{rd:ode}
\begin{bmatrix} \mathbf{f}_0 + \mathbf{f}_1\\
                \mathbf{f}_2 + \mathbf{f}_3\\
                \mathbf{f}_a \end{bmatrix} = \mathbf{0}
\end{align}

The naive approach to linearizing these equations of motion would be to solve
for the ($\dot{q}_i$, $\dot{u}_i$) terms to construct the right-hand side of
\begin{align}
\begin{bmatrix}\dot{\mathbf{q}} \\ \dot{\mathbf{u}}\end{bmatrix} =
    \mathbf{f}(\mathbf{q}, \mathbf{u}, t)
\end{align}
and then take the jacobian of $\mathbf{f}$ with respect to $[\mathbf{q}
\quad \mathbf{u}]^T$. However, this will give the incorrect results. As
discussed in Schwab, this system has a critical speed where all the eigenvalues
are 0 under the following conditions: the system is linearized about the
upright rolling steady-state configuration, and all parameters ($r$, $m$, and
$g$) are unity. When these conditions are met, the critical speed is
$v=\frac{1}{\sqrt{3}}$.

When following the naive approach, 8 of the eigenvalues are 0 and 4 are
non-zero under the same conditions:
\begin{align}
\sqrt{-v^2},\quad -\sqrt{-v^2},\quad \sqrt{-\frac{12}{5} v^2 +
\frac{4}{5}},\quad -\sqrt{-\frac{12}{5} v^2 + \frac{4}{5}}
\end{align}
At the critical speed, two of these non-zero eigenvalues go to zero. But at any
forward speed, there is a nonzero imaginary eigenvalue pair. Having illustrated
the failings of the simple approach to linearization of a constrained multibody
system defined with Kane's method, we will now outline a linearization
procedure which addresses this issue and then revisit this example applying the
proper linearization procedure.
